---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# blsr

An R package for obtaining and cleaning data from the Bureau of Labor Statistics (BLS). This package uses the BLS API. 

## Highlights

- Simple-to-use functions for downloading data from the most commonly used BLS databases. 

- Cleans the data returned by the BLS API and converts the data into a panel format. 

- Supplements the blscrapeR package by converting inputs to BLS series ID strings.

- Impelements error checks for BLS database idiosyncracies.  

## Usage

A detailed usage description can be found in the [vignette](https://github.com/davidsovich/blsr/blob/master/vignettes/blsr.pdf). 

This package currently supports the following BLS databases:

- Current employment statistics (CES)

- Job openings and labor turnover survey (JOLTS)

- Local area unemployment statistics (LAUS)

In the near future, this package should also support these additional BLS databases:

- Quarterly Census of Employment and Wages (QCEW)

- Current Population Survey (CPS)

- Consumer Price Index (CPI)

- Current Expenditure Survey (CE)

Examples:

```{r, eval = FALSE}
library(blscrapeR)


#Download data from CES

   # Custom CES state data series
   ces_df = ces_download(
      bls_key = Sys.getenv("BLS_KEY"),
      start_year = 2010,
      end_year = 2015,
      adjustment = "U",
      industries = "05000000",
      data_types = c("01", "03", "11"),
      states = "1900000"
   )
   
   # Pre-built CES seasonally adjusted non-farm payroll series
   ces_df = ces_employment(
      bls_key = Sys.getenv("BLS_KEY"),
      series = "nfp",
      start_year = 2010,
      end_year = 2015,
      adjustment = "S"
   )
   
   # Pre-built CES non-seasonally adjusted sector series
   ces_df = ces_employment(
      bls_key = Sys.getenv("BLS_KEY"),
      series = "super",
      start_year = 2010,
      end_year = 2018,
      adjustment = "U"
   )

#Download data from JOLTS

   # Custom JOLTS hire and quit rates and levels series
   jolts_df = jolts_download(
      bls_key = Sys.getenv("BLS_KEY"),
      start_year = 2010,
      end_year = 2015,
      adjustment = "S",
      industries = "000000",
      data_types = c("HI", "QU"),
      data_levels = c("R", "L")
   )
   
   # Pre-built JOLTS seasonally adjusted non-farm hires series
   jolts_df = jolts_hires(
      bls_key = Sys.getenv("BLS_KEY"),
      series = "nfp",
      start_year = 2010,
      end_year = 2015,
      adjustment = "S"
   )
   
   # Pre-built JOLTS non-seasonally adjusted sector separations series
   jolts_df = jolts_seps(
      bls_key = Sys.getenv("BLS_KEY"),
      series = "super",
      start_year = 2010,
      end_year = 2015,
      adjustment = "S"
   )


# Local area unemployment statistics

   # Custom LAUS data series
   laus_df = laus_download(
      bls_key = Sys.getenv("BLS_KEY"),
      start_year = 2010,
      end_year = 2015,
      adjustment = "S", 
      states = c("ST0100000000000", "ST0200000000000"), 
      data_types = c("03")
   )
   
   # Pre-built LAUS seasonally adjusted unemployment rate series
   laus_df = laus_urate(
      bls_key = Sys.getenv("BLS_KEY"),
      start_year = 2010,
      end_year = 2015,
      adjustment = "S"
   )

```

## Installation

The blsr package is not available on CRAN. You can install the development version from Github:

```{r, eval = FALSE}
library("devtools")
devtools::install_github("davidsovich/blsr")
```

## Contact

dsovich `AT` uky.edu

## History

- March 22, 2019: Developmental release

- December 16, 2020: Partial update to incorporate changes to JOLTS series string formats. 
