---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# blsr

An R package for obtaining and cleaning data from the Bureau of Labor Statistics (BLS). This product uses the BLS API and the blscrapeR package.  

## Highlights

- Simple-to-use functions for downloading data from the most commonly used BLS databases. 

- Cleans the data returned by the BLS API and converts the data into a panel format. 

- Converts user inputs into BLS series ID strings and implements error-checks. 

## Usage

A more detailed usage description can be found in the [vignette](www.davidsovich.com). 

Current, the supported BLS databases include:

- Current employment statistics (CES)

- Job openings and labor turnover survey (JOLTS)

- Local area unemployment statistics (LAUS)

Examples:

```{r, eval = FALSE}
library(blscrapeR)
#Download data from CES

#Download data from JOLTS

```

## Installation

The blsr package is not available on CRAN. You can install the development version from Github:

```{r, eval = FALSE}
library("devtools")
devtools::install_github("davidsovich/blsr")
```

## History

- March 22, 2019: Developmental release
